---
title: "2nd Graph"
format: html
editor: visual
---

## List of countries and dependencies by population\[3\]

The graph shows a list of countries and dependencies by population in 2018. This graph and dataset is from a [Wikipedia page](https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population#References).

![Original Visualization \[2\]](mapimage.png){fig-align="center"}

While the graph title says that the data is from 2018, there is updated data available in a table format on the same page.

Each square in the graph represents 500,000 people. The graph shows each country's population and percentage of the world along with a few states and cities. This map has also calculated the continent-wise population. The sizes are different than that of a regular map, it allows for easy comparison of population sizes around the world. You can quickly see how the population of **India** compares with, for example, **Nigeria** or **Brazil**, as the size of the country on the map is proportional to its population. This approach utilizes the available space well by resizing countries according to their population rather than land area. This way, smaller countries with large populations (e.g., Bangladesh, Japan) are still emphasized rather than getting lost in a traditional map.

#### Flaws :

1.  Since there is no proper outline for each country, some colours are harder to distinguish from each other.

<!-- -->

2.  Each country's colour has been assigned randomly and not according to their population sizes.
3.  The graph looks way too crowded since the creator put the population size, percentage of world as well as the name of the country all on the map.
4.  The square boxes representing 500,000 people are not actually needed, it also makes the map look glitchy and non-aesthetic.

To overcome these visualization flaws and to make it most user friendly, here are the redesigns.

## Redesign Visualization 1.

```{r,warning=FALSE, message=FALSE}
#| echo: false
#install all the required libraries
library(leaflet)
library(dplyr)
library(readr)
library(sf)
library(RColorBrewer)
library(htmltools)

#reading the CSV file
data <- read_csv("countries_population.csv",show_col_types = FALSE)

#Cleaning the CSV file
data$Location <- gsub("_", " ", data$Location)

# --- Clean country names so they match the map's names ---
data$map_name <- trimws(sub("\\([^()]*\\)$", "", data$Location))   # drop "(UK)", "(US)" etc.
fix <- c("United States" = "USA", "United Kingdom" = "UK",
         "Republic of the Congo" = "Republic of Congo", "East Timor" = "Timor-Leste",
         "Eswatini" = "Swaziland", "Vatican City" = "Vatican", "Curasao" = "Curacao",
         "U.S. Virgin Islands" = "Virgin Islands, US",
         "British Virgin Islands" = "Virgin Islands, British",
         "Cocos (Keeling) Islands" = "Cocos Islands")
data$map_name <- ifelse(data$map_name %in% names(fix), fix[data$map_name], data$map_name)
# Some countries are split into separate islands on the map
split <- data.frame(
  ID = c("Trinidad", "Tobago", "Antigua", "Barbuda", "Saint Kitts", "Nevis",
         "Saint Vincent", "Grenadines", "Saint Helena", "Ascension Island"),
  map_name = rep(c("Trinidad and Tobago", "Antigua and Barbuda", "Saint Kitts and Nevis",
                   "Saint Vincent and the Grenadines",
                   "Saint Helena, Ascension and Tristan da Cunha"), each = 2))

#Creating the intervals
bins <- c(0, 100000, 200000, 300000, 500000,
          3000000, 5000000, 9000000, 11000000,
          16000000, 20000000, 25000000, 40000000,
          60000000, 67000000, 90000000, 200000000,
          400000000, 1500000001)
#Legend 
labels <- c(
  "0 - 100,000",
  "100,001 - 200,000",
  "200,001 - 300,000",
  "300,001 - 500,000",
  "500,001 - 3,000,000",
  "3,000,001 - 5,000,000",
  "5,000,001 - 9,000,000",
  "9,000,001 - 11,000,000",
  "11,000,001 - 16,000,000",
  "16,000,001 - 20,000,000",
  "20,000,001 - 25,000,000",
  "25,000,001 - 40,000,000",
  "40,000,001 - 60,000,000",
  "60,000,001 - 67,000,000",
  "67,000,001 - 90,000,000",
  "90,000,001 - 200,000,000",
  "200,000,001 - 400,000,000",
  "400,000,001 - 1,500,000,001"
)

data$population_group <- cut(data$Population, breaks = bins,
                             include.lowest = TRUE,
                             labels = labels)

# We need colours for the number of intervels we have
num_colors <- length(labels)

color_palette <- colorRampPalette(brewer.pal(9, "Blues"))  
colors <- color_palette(num_colors)

#Importing the world dataset
world <- st_as_sf(maps::map("world", fill = TRUE, plot = FALSE))

#Joining the map shapes to the population data using the cleaned names
world_data <- world %>%
  left_join(split, by = "ID") %>%
  mutate(map_name = coalesce(map_name, ID)) %>%
  left_join(data, by = "map_name") %>%
  mutate(color = colors[as.numeric(cut(Population, breaks = bins, include.lowest = TRUE))])


# Visual 1
leaflet_map <- leaflet(world_data) %>%
  addTiles() %>%
  setView(lng = 0, lat = 20, zoom = 2) %>%
  addPolygons(fillColor = ~color,
              weight = 1,
              opacity = 1,
              color = 'white',
              dashArray = '3',
              fillOpacity = 0.7,
              highlightOptions = highlightOptions(
                weight = 5,
                color = '#666',
                dashArray = '',
                fillOpacity = 0.7,
                bringToFront = TRUE),
              label = ~lapply(ifelse(is.na(Population),
                                     paste0("<b>", ID, "</b><br>No population data"),
                                     paste0("<b>", Location, "</b><br>",
                                            "Population: ", format(Population, big.mark = ",", scientific = FALSE), "<br>",
                                            "% of World: ", `%_ofworld`)),
                              htmltools::HTML),
              labelOptions = labelOptions(style = list('font-weight' = 'normal', 
                                                       padding = '3px 8px'), 
                                          textsize = "15px", 
                                          direction = "auto")) %>%
  addLegend(pal = colorFactor(colors, NULL), 
            values = ~population_group, 
            labels = labels,  # Update labels to show intervals
            opacity = 0.7, 
            title = "Population Groups", position = "bottomright")

leaflet_map

```

**Improvements:**

1.  Each country is outlined with white break lines, this shows clear separation between countries.

2.  This is an interactive map, when you zoom you can seen all the names of the country, state, cities.

    This can be helpful later on for some analysis.

3.  When you hover over the map it shows you the name of the country, the total population there and the percent of world it makes up. This information is not visible before hovering on that country, this keeps the map less cluttered.

4.  The colours are in ombre dark blue to white shade, where dark blue are the densely populated countries and as the colour goes towards white those are lesser populated areas.

5.  The colours are given based on certain intervals which I chose by looking at the dataset.

6.  The sizes of each country is true to the real map size so that people won't get confused by seeing offset sizes.

One piece of information that this redesign does not have is the population of continents. To overcome this flaw we can use a Drilldown chart.

## Redesign Visualization 2. \[4\]

This graph starts out with a bar graph of the total population of each continent.

1.  It gives us the data we were missing out on by doing just maps.
2.  This is also an interactive graph: click any continent and the chart drills down to show the population of every country in that continent.
3.  Both levels are arranged in descending order.
4.  The bar graph type visualization make it very easy to notice which continent has the most population , which countries are more populated ect with just a look at the visualization.

```{r,message=FALSE, warning=FALSE}
#| echo: false
library(dplyr)
library(purrr)
library(highcharter)

data1 <- read.csv("countries_population_continents.csv") %>%
  filter(Location != "World") %>%                      # keep countries only
  mutate(Continent = ifelse(Continent == "Australia", "Oceania", Continent))

# Total population for each continent, largest first
by_continent <- data1 %>%
  group_by(Continent) %>%
  summarise(Population = sum(Population, na.rm = TRUE), .groups = 'drop') %>%
  arrange(desc(Population))

# Drilldown data for every continent: its countries, largest first
drilldown_series <- data1 %>%
  group_by(Continent) %>%
  group_split() %>%
  map(function(df) list(
    id = df$Continent[1],
    name = df$Continent[1],
    data = df %>%
      arrange(desc(Population)) %>%
      transmute(name = Location, y = Population) %>%
      list_parse()
  ))

highchart() %>%
  hc_chart(type = "column") %>%
  hc_title(text = "Population Drilldown by Continent and Country") %>%
  hc_subtitle(text = "Click a continent to see its countries") %>%
  hc_xAxis(type = "category") %>%
  hc_yAxis(title = list(text = "Population")) %>%
  hc_tooltip(pointFormat = "<b>{point.y:,.0f}</b> people") %>%
  hc_legend(enabled = FALSE) %>%
  hc_add_series(
    name = "Total Population",
    colorByPoint = TRUE,
    data = by_continent %>%
      transmute(name = Continent, y = Population, drilldown = Continent)
  ) %>%
  hc_drilldown(series = drilldown_series)

```

**Special efforts:**

The data set was web-scrapped from wikipedia page table. After the data was scrapped, it was cleaned by removing spaces and '\_'.

Some letters had to be rewritten manually since accented letters were not recognized.

Created interactive plots to enhance customer/user experience.

Used a website to learn how to do Drilldown charts.

Had to mutate the dataset by mapping continents to the different countries for the second visualization
